---
title: "fixing_drift"
output: html_document
---

# Introduction

This vignette demonstrates the complete workflow for drift correction on optical
water quality sensor data using field notes to linearly correct drift.

# Prepare Environment
```{r setup}
knitr::opts_chunk$set(
  eval = FALSE,
  echo = TRUE
)
```

# Load libraries
```{r}
#Installing and loading all packages
invisible(
  lapply(c(
    "tidyverse", # Data manipulation
    "janitor", # Clean dirty data
    "lubridate", # Date-Time Manipulation
    "rvest", # HTML Retrieval and Manipulation
    "readxl", # Reading excel files
    "here", # Easy, cross platform file referencing
    "ggplot2", # Plotting libraries
    "ggpubr",
    "plotly",
    "devtools", # For downloading GitHub packages
    "remotes",
    "yaml",
    "ggpubr"
  ),
  function(x) {
    if (x %in% installed.packages()) {
      suppressMessages({
        library(x, character.only = TRUE)
      })
    } else {
      suppressMessages({
        install.packages(x)
        library(x, character.only = TRUE)
      })
    }
  })
)

load_all()
```

# Load some data in to play with

```{r}
# Read in 2023, 2024, and 2025 sensor data
sensor_data <- read_rds(here("..", "poudre_sonde_network", "data", "manual_data_verification", "complete_dataset", "23_25_hv_pull"))
```

```{r}
# Flatten the nested list and combine all tibbles
data_list <- sensor_data %>%
  map_dfr(~ map_dfr(.x, bind_rows)) %>%
  filter(
    parameter %in% c("Turbidity", "Chl-a Fluorescence")
  ) %>%
  split(f = list(.$site, .$parameter), sep = "-") %>%
  discard(~ is.null(.x) | nrow(.x) == 0)
```


# Read in the field notes to join it with the data.

```{r}
mw_notes <- load_mWater(creds = read_yaml(here("..", "poudre_sonde_network", "creds", "mWaterCreds.yml")))

field_notes <- ross.wq.tools::grab_mWater_sensor_notes(mw_notes)
malfunction_notes <- ross.wq.tools::grab_mWater_malfunction_notes(mw_notes)
```

# Join the field notes with the sensor data
```{r}
notes_list <- data_list %>%
  map(function(df) {
    ross.wq.tools::add_field_notes(
      df = df, 
      notes = field_notes
    )
  })
```

# Apply the malfunction flag and the drift flag 

I want to see how well these functions are working independently and with each other,
without relying on the network check.

```{r}
flagged_list <- notes_list %>%
  map(function(df) {
    df %>%
      add_field_flag() %>%
      add_malfunction_flag(malfunction_records = malfunction_notes) %>%
      add_drift_flag()
  })
```

```{r}
# Create a combined column for all drift/site_visit combinations
flagged_list <- flagged_list %>%
  map(function(df) {
    df %>%
      mutate(
        drift = ifelse(grepl("drift", flag), TRUE, FALSE),
        site_visit = ifelse(grepl("site visit|sv window", flag), TRUE, FALSE),
        # Create combined factor for plotting
        drift_visit_combo = case_when(
          !drift & !site_visit ~ "No drift, No visit",
          drift & !site_visit ~ "Drift only", 
          !drift & site_visit ~ "Site visit only",
          drift & site_visit ~ "Drift + Site visit",
          TRUE ~ "Other"  # fallback
        )
      )
  })
```

# Make plotting functions to easily look at different sections of the data
```{r}
# Updated plotting function
plot_drift <- function(df) {
  ggplot(data = df, aes(x = DT_round, y = mean, color = drift_visit_combo)) +
    geom_point(aes(alpha = drift_visit_combo)) +
    scale_color_manual(values = c(
      "No flags" = "grey",
      "Drift only" = "tomato", 
      # "Site visit only" = "blue",
      "Drift + Site visit" = "purple"
    )) +
    scale_alpha_manual(values = c(
      # "No drift, No visit" = 0.3,
      "Drift only" = 0.8,
      # "Site visit only" = 0.6, 
      "Drift + Site visit" = 1.0
    )) +
    labs(color = "Status", alpha = "Status")
}
```

# Clean up the environment

```{r}
rm(
  # HV sensor data
  sensor_data, data_list,
  # MW notes data
  mw_notes, field_notes, malfunction_notes,
  # Derived combined HV and MW data
  notes_list
)

gc()
```

# >>>> Detecting drift issues <<<<
```{r}

```


# >>>> Drafting drift solutions <<<<

## Field notes implementation

### Analyzing the differences between pre and post calibration data

The assumption here is that the data in field notes should be enough to do a 
one point drift correction. The way that one point drift corrections work (at 
least using the `{driftr}` methods) is that a difference between two data points 
is taken, and then is applied to to the data in a temporally weighted way. When
we tried to fix this with the calibration data we fell very short because the differences
between those two data points was often very small, often less than ~3 NTU, while
the difference that we were trying to correct for in the drift was often around 
~100 NTU. This method did not work then, but the assumption here is that the
pre and post calibration data points captured in mWater field notes should be enough
to correct this drift correction. Here I want to analyse these differences to get 
a sense if this will work or not. 

I am subtracting the post calibration number from the pre calibration number.
Under the assumption that the pre calibration number should typically be bigger
than the post calibration number, we should expect mostly negative values

#### Plot Histograms of the differences between these values to see whats going on.

There are some outliers that make this analysis tough, so for now lets remove 
those, and plot what the histogram looks like then.

```{r}
diff_df <- flagged_list %>% 
  bind_rows() %>% 
  filter(parameter == "Turbidity", 
         !is.na(turb_pre_clean) & (!is.na(turb_post_cal) | !is.na(turb_post_clean))) %>% 
  mutate(
    diff_clean = turb_post_clean - turb_pre_clean,
    diff_cal = turb_post_cal - turb_pre_clean
  ) %>% 
  filter(
    diff_clean > -25 & diff_clean < 25,
    diff_cal > -25 & diff_cal < 25
  )

# Create histograms for cleaning effect (pre_clean to post_clean)
clean_plot <- ggplot(data = diff_df) +
  geom_histogram(aes(x = diff_clean), bins = 30, fill = "steelblue") +
  labs(title = "Effect of Cleaning", 
       x = "Post-Clean - Pre-Clean (NTU)") +
  theme_minimal()

# Create histograms for calibration effect (pre_clean to post_cal)
cal_plot <- ggplot(data = diff_df) +
  geom_histogram(aes(x = diff_cal), bins = 30, fill = "darkgreen") +
  labs(title = "Effect of Calibration", 
       x = "Post-Cal - Pre-Clean (NTU)") +
  theme_minimal()

# Compare them side by side
ggarrange(clean_plot, cal_plot, ncol = 2)
```

The fact that these numbers are not bigger is concerning to me. 

### Example of drift with field notes that can be used to fix it.
```{r}
# Example drift with field notes
# og_example_drift_fn <- flagged_list$`cottonwood-Turbidity` %>% 
#   filter(between(DT_round, as.POSIXct("2025-06-25"), as.POSIXct("2025-07-07")))

# Updated plotting function
plot_drift <- function(df) {
  ggplot(data = df, aes(x = DT_round, y = mean, color = drift_visit_combo)) +
    geom_point(aes(alpha = drift_visit_combo)) +
    scale_color_manual(values = c(
      "No drift, No visit" = "grey",
      "Drift only" = "tomato", 
      "Site visit only" = "blue",
      "Drift + Site visit" = "purple"
    )) +
    scale_alpha_manual(values = c(
      "No drift, No visit" = 0.3,
      "Drift only" = 0.8,
      "Site visit only" = 0.6,
      "Drift + Site visit" = 1.0
    )) +
    labs(
      color = "Status", 
      alpha = "Status",
      x = "Datetime",
      y = "Turbidity"
    ) +
    theme_bw()
}

example_drift_fn <- flagged_list$`cottonwood-Turbidity` %>% 
  filter(between(DT_round, as.POSIXct("2025-06-25"), as.POSIXct("2025-07-07"))) %>% 
  mutate(
    # Fill in the missed drift:
    drift = ifelse(between(DT_round, as.POSIXct("2025-06-30"), as.POSIXct("2025-07-03")), TRUE, drift),
    # Create combined factor for plotting
    drift_visit_combo = case_when(
      !drift & !site_visit ~ "No drift, No visit",
      drift & !site_visit ~ "Drift only", 
      !drift & site_visit ~ "Site visit only",
      drift & site_visit ~ "Drift + Site visit",
      TRUE ~ "Other"  # fallback
    )
  )

plot_drift(example_drift_fn) +
  ylim(0, 100)
```

Here the `add_drift_flag()` function was able to detect drift. There is some 
weirdness going on in the middle of the flagged data, but for now we are going
to ignore that to keep making progress on drift correction. We will improve 
drift detection later. For now, we are just going to tweak it so that all of those
missed points also have drift in them, to test the method

Also note the difference in NTU from the end of drift to the new site visit group. 
It's close to 100NTU.

Here is the plan to correct drift:

1. Identify "drift groups". Those consecutive groups of data all belong to a single drift instance.
2. Find the appropriate "before" and "after" numbers to do a one point drift correction.
3. Apply the appropriate one point drift correction information to the right drift group.

We do steps 1 and 2 in reverse order. First, we find "site visit groups", and then
look for "drift groups" within those groups. The assumption here is that each 
site visit group _should_ have the information that we need to correct the single
drift group inside of it. Just to make these assumptions clear:

- Each site visit group has one drift group.
- Each site visit's drift correction information (pre, post calibration info) is appropriate for the related drift group.

These assumptions are not always met with our current detection and correction methods.
Whatever, we ball. Lets just keep going and then build from what we find out.

```{r}
# install.packages("groupdata2")
# library(groupdata2)

# install.packages("forecast")
# library(forecast)

# lets make this a function and see what happens
drift_corrector <- function(df) {
  
  corrected_drift_df <- df %>% 
    # there are `last_site_visit` groups that we can use, but the information for a drift correction
    # for one of these groups is actually in the following group's (much like the calibration correction).
    
    # Fortunately, we can fill backwards on these groups in order to grab the 
    # relevant information for the groups that we want to correct. For now,
    # lets only fill backwards the columns that we know we are going to need.
    
    tidyr::fill(turb_pre_clean, .direction = "up") %>% 
    tidyr::fill(turb_post_clean, .direction = "up") %>% # Use this if there is no post cal value
    tidyr::fill(turb_post_cal, .direction = "up") %>% # Use this preferentially
    
    # Character mutate of last site visit so that groupdata2 plays ball with us
    mutate(last_site_visit_chr = as.character(last_site_visit)) %>% 
    
    # Now that we have the information that we need from each site visit group,
    # lets split them up
    groupdata2::splt(
      n = "auto",
      method = "l_starts",
      starts_col = "last_site_visit_chr"
    ) %>% 
    map(function(sv_group_df){
      
      # applying a mutate here to work with `group_factor()`, but this mutate should
      # actually be made when I "create a combined column for all drift/site_visit 
      # combinations"
      
      # Drift = 1, NO Drift = 0
      df <- sv_group_df %>% 
        mutate(drift = ifelse(drift == TRUE, 1, 0)) 
      
      # Not all of these dfs have drift, so we can just skip those
      # Check if there is only 1 group, and if there is, check if it is drift,
      # else skip. If there are 2 groups, one of them is drift, so we are not going to check 
      # that right now
      if(length(unique(df$drift)) == 1 && unique(df$drift) == 0) return(df)
      
      # cal one point drift handles missing calibration data, so lets split these 
      # Site visit groups into drift and non drift groups
      df <- df %>% 
        groupdata2::splt(
          n = "auto",
          method = "l_starts",
          starts_col = "drift"
        ) %>% 
        map(
          function(drift_group) {
            # First check if it is not a drift group, and skip if not
            if(unique(drift_group$drift) == 0) return(drift_group)
            
            # >>>> DRIFT CORRECTION STEP <<<<
            # If it is a drift group, pass it through the weighting and 
            # one point calibration functions
            
            # Check which "post" value we should use, and then use that.
            post_cal <- df[["turb_post_cal"]][[nrow(df)]]
            post_clean <- df[["turb_post_clean"]][[nrow(df)]]
            
            if (!is.na(post_cal)) {
              corrected_drift_group <- drift_group %>% 
                ross.wq.tools::cal_wt(dt_col = "DT_round") %>% 
                cal_exp_one_point_drift(
                  lm_trans_col = "mean", 
                  pre_col = "turb_pre_clean", 
                  post_col = "turb_post_cal", 
                  wt_col = "wt",
                  drift_type = "exponential",
                  correction_type = "multiplicative"
                )
            } else if (!is.na(post_clean)) {
              corrected_drift_group <- drift_group %>% 
                ross.wq.tools::cal_wt(dt_col = "DT_round") %>% 
                cal_exp_one_point_drift(
                  lm_trans_col = "mean", 
                  pre_col = "turb_pre_clean", 
                  post_col = "turb_post_clean", 
                  wt_col = "wt",
                  drift_type = "exponential",
                  correction_type = "multiplicative"
                )
            } else {
              return(drift_group)
            }
            
            return(corrected_drift_group)
          }
        ) %>% 
        reduce(full_join)
    }) %>% 
    reduce(full_join)
  
  return(corrected_drift_df)
}

```

```{r}
ex_dr_fn_dg <- drift_corrector(example_drift_fn)
```

Now that we have corrected drift, lets see what happened:
```{r}
ggplot(ex_dr_fn_dg, aes(x = DT_round)) +
  geom_point(aes(y = mean, color = "Uncorrected"), alpha = 0.7) +
  geom_point(aes(y = mean_drift_trans, color = "Drift Correction"), alpha = 0.7) +
  labs(
    x = "Datetime",
    y = "Turbidity",
    color = "Data Type"
  ) +
  scale_alpha_manual(values = c("Uncorrected" = 0.3,
                                "Drift Correction" = 0.6
  )) +
  scale_color_manual(values = c("Uncorrected" = "gray50", 
                                "Drift Correction" = "steelblue")
  ) +
  theme_bw() 
```

Looks pretty good. There is a hump there that might suggest that the drift is
slightly faster than exponential growth, but overall I would say the function works.

This example I was very careful with the data and took good care to make sure 
everything was good so that it would work. The plan is to apply this liberally
across a bunch of data. That is when I expect this to break. 

Lets try to break it right now, even.

```{r}
example_drift_fn_2 <- flagged_list$`cottonwood-Turbidity` %>% 
  filter(between(DT_round, as.POSIXct("2025-05-01"), as.POSIXct("2025-05-30"))) 

plot_drift(example_drift_fn_2) +
  ylim(0, 100) 
```

This data, which comes from the same data set that I just used to correct the drift
correctly, breaks a lot of the assumptions that I was talking about earlier. Lets 
see what happens when we apply this function willy nilly.

```{r}
ex_dr_fn_dg_2 <- drift_corrector(example_drift_fn_2)

ggplot(ex_dr_fn_dg_2, aes(x = DT_round)) +
  geom_line(aes(y = mean, color = "Uncorrected"), alpha = 0.7) +
  geom_line(aes(y = mean_drift_trans, color = "One-Point Correction"), alpha = 0.7) +
  geom_point(data = filter(ex_dr_fn_dg_2, site_visit == TRUE), 
             aes(y = mean), color = "black", size = 2) +
  labs(
    title = "Turbidity: Drift Corrected vs Uncorrected",
    subtitle = "Cottonwood Site",
    x = "Date",
    y = "Turbidity (NTU)",
    color = "Data Type"
  ) +
  scale_color_manual(values = c("Uncorrected" = "gray50", 
                                "One-Point Correction" = "steelblue")
  ) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  ylim(0, 100)
```

Yeah, that's no good. A problem for later. Moving on.

## Data heuristic implementation

### Example of drift with no field notes that can be used to fix it. 

```{r}
example_drift_dh <- flagged_list$`riverbend-Turbidity` %>% 
  filter(between(DT_round, as.POSIXct("2023-04-20"), as.POSIXct("2023-04-30")))

plot_drift(example_drift_dh) +
  # xlim(as.POSIXct("2023-04-20"), as.POSIXct("2023-04-30")) +
  ylim(0, 100)
```

Before, we were able to use the information in the field notes to fix the problem. 
We will not always have this at hand for older data, and there may be drift with
no site visit information as well. So lets see what we can do in these weirder 
situations. 

First, I am going to use the example that I used before, because the drift detection
was too greedy and flagged more data than there should be here. 

The overall idea here should be basically the same. We need to find a value that 
represents the peak of the drift, and a value that represents the data when it 
came back down. The trouble here is figuring out how to assimilate this into the 
code that we already have. 

For now lets ignore site visits and see what we can do with just drift group
information.

```{r}
# example_drift_fn

plot_drift(example_drift_fn) +
  ylim(0, 100)
```

```{r}
# HERE
# I left off trying to make the drift correctin function work for two data points that are not in columns
# There is another, bigger issue of accurate drift detection, or rather ending the drift detection algorithm 
# at the right time, but that is something for later.

corrected_drift_df <- example_drift_fn %>% 
  # Lets split this data up by drift groups
  groupdata2::group(
    n = "auto",
    method = "l_starts",
    starts_col = "drift"
  )
map(function(drift_group_df){
  
  # applying a mutate here to work with `group_factor()`, but this mutate should
  # actually be made when I "create a combined column for all drift/site_visit 
  # combinations"
  
  # Drift = 1, NO Drift = 0
  df <- drift_group_df %>% 
    mutate(drift = ifelse(drift == TRUE, 1, 0)) 
  
  # Not all of these dfs have drift, so we can just skip those
  # Check if there is only 1 group, and if there is, check if it is drift,
  # else skip. If there are 2 groups, one of them is drift, so we are not going to check 
  # that right now
  if(length(unique(df$drift)) == 1 && unique(df$drift) == 0) return(df)
  
  # cal one point drift handles missing calibration data, so lets split these 
  # Site visit groups into drift and non drift groups
  df <- df %>% 
    groupdata2::splt(
      n = "auto",
      method = "l_starts",
      starts_col = "drift"
    ) %>% 
    map(
      function(drift_group) {
        # First check if it is not a drift group, and skip if not
        if(unique(drift_group$drift) == 0) return(drift_group)
        
        # >>>> DRIFT CORRECTION STEP <<<<
        # If it is a drift group, pass it through the weighting and 
        # one point calibration functions
        
        # Check which "post" value we should use, and then use that.
        post_cal <- df[["turb_post_cal"]][[nrow(df)]]
        post_clean <- df[["turb_post_clean"]][[nrow(df)]]
        
        if (!is.na(post_cal)) {
          corrected_drift_group <- drift_group %>% 
            ross.wq.tools::cal_wt(dt_col = "DT_round") %>% 
            cal_exp_one_point_drift(
              lm_trans_col = "mean", 
              pre_col = "turb_pre_clean", 
              post_col = "turb_post_cal", 
              wt_col = "wt",
              drift_type = "exponential",
              correction_type = "multiplicative"
            )
        } else if (!is.na(post_clean)) {
          corrected_drift_group <- drift_group %>% 
            ross.wq.tools::cal_wt(dt_col = "DT_round") %>% 
            cal_exp_one_point_drift(
              lm_trans_col = "mean", 
              pre_col = "turb_pre_clean", 
              post_col = "turb_post_clean", 
              wt_col = "wt",
              drift_type = "exponential",
              correction_type = "multiplicative"
            )
        } else {
          return(drift_group)
        }
        
        return(corrected_drift_group)
      }
    ) %>% 
    reduce(full_join)
}) %>% 
  reduce(full_join)


```

